### 2022-04-09 outputs/paper/paper.Rmd
Initial commit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
data <- read.csv("cleaned.csv", header=T)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
include_graphics(here::here("inputs/literature/Figure Tanz.png"))
```

Talk more about it.

```{r}
data |> 
  select(region, no_mass_media_f, no_mass_media_m, all_f, all_m, number_f, number_m) |> 
  #slice(1:10) |> 
  kable(
    caption = "First ten rows of a dataset of factors that contribute to commonality of marriage",
    col.names = c("Region", "No Media Female", "No Media Male",
                  "All Media Female", "All Media Male", "Total Number of Female",
                  "Total Number of Male"),
    digits = 1,
    booktabs = TRUE, 
    linesep = "")
```

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r}
data |> 
  select(-X)|>
  datasummary_skim(type = "numeric",
                   title = "Summary of numerical ditrubution indicators 
                   variables for regions")
```

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=20, fig.height=4}

fm<- 
  data |>
  #slice(1:10) |> 
  ggplot(mapping = aes(x = reorder(region, -no_mass_media_f), y = no_mass_media_f, fill=region)) +
  geom_bar(stat = "identity", show.legend = FALSE)+
  labs(x = "Region", y = "Females Without Mass Media")+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE)
                     #option = "magma")

m<- 
  data |>
  #slice(1:10) |> 
  ggplot(mapping = aes(x = reorder(region, -no_mass_media_m), y = no_mass_media_m, fill=region)) +
  geom_bar(stat = "identity", show.legend = FALSE)+
  labs(x = "Region", y = "Males Without Mass Media")+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE)

fm/m

```

Talk way more about it.

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=20, fig.height=4}

fm2<- 
  data |>
  #slice(1:10) |> 
  ggplot(mapping = aes(x = reorder(region, -all_f), y = all_f, fill=region)) +
  geom_bar(stat = "identity", show.legend = FALSE)+
  labs(x = "Region", y = "Females With All 3 Mass Media")+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE,
                     option = "magma")

m2<- 
  data |>
  #slice(1:10) |> 
  ggplot(mapping = aes(x = reorder(region, -all_m), y = all_m, fill=region)) +
  geom_bar(stat = "identity", show.legend = FALSE)+
  labs(x = "Region", y = "Males With All 3 Mass Media")+
  theme_classic() +
  scale_fill_viridis(discrete = TRUE,
                     option = "magma")

fm2/m2

```

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=10, fig.height=10}

newsm<- 
data |>
  ggplot(mapping = aes(x = all_m, y = read_news_weekly_m, color = region)) +
  geom_point()+
  labs(x = "All Males With Media", y = "Read Newspaper Weekly")+
  theme(legend.position="top")+
  scale_colour_viridis_d()

telem<- 
data |>
  ggplot(mapping = aes(x = all_m, y = watch_tele_weekly_m, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "All Males With Media", y = "Watch Television Weekly")+
  scale_colour_viridis_d()

radm<- 
data |>
  ggplot(mapping = aes(x = all_m, y = listen_radio_weekly_m, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "All Males With Media", y = "Listen to Radio Weekly ")+
  scale_colour_viridis_d()

newsm/telem/radm

```

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=10, fig.height=10}

newsf<- 
data |>
  ggplot(mapping = aes(x = all_f, y = read_news_weekly_f, color = region)) +
  geom_point()+
  labs(x = "All Females With Media", y = "Read Newspaper Weekly")+
  theme(legend.position="top")+
  scale_colour_viridis_d(option = "plasma")

telef<- 
data |>
  ggplot(mapping = aes(x = all_f, y = watch_tele_weekly_f, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "All Females With Media", y = "Watch Television Weekly")+
  scale_colour_viridis_d(option = "plasma")

radf<- 
data |>
  ggplot(mapping = aes(x = all_f, y = listen_radio_weekly_f, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "All Females With Media", y = "Listen to Radio Weekly ")+
  scale_colour_viridis_d(option = "plasma")

newsf/telef/radf

```

```{r}

rwf<- 
data |>
  ggplot(mapping = aes(x = listen_radio_weekly_f, y = watch_tele_weekly_f, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(title= "Females", x = "Listen to Radio Weekly", y = "Watch Television Weekly")

rnf<- 
data |>
  ggplot(mapping = aes(x = listen_radio_weekly_f, y = read_news_weekly_f, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "Listen to Radio Weekly", y = "Read Newspaper Weekly")

nwf<- 
data |>
  ggplot(mapping = aes(x = read_news_weekly_f, y = watch_tele_weekly_f, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "Read Newspaper Weekly", y = "Watch Television Weekly")
rwf/rnf/nwf

```

```{r}

rwm<- 
data |>
  ggplot(mapping = aes(x = listen_radio_weekly_m, y = watch_tele_weekly_m, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(title= "Males", x = "Listen to Radio Weekly", y = "Watch Television Weekly")

rnm<-
data |>
  ggplot(mapping = aes(x = listen_radio_weekly_m, y = read_news_weekly_m, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "Listen to Radio Weekly", y = "Read Newspaper Weekly")

nwm<- 
data |>
  ggplot(mapping = aes(x = read_news_weekly_m, y = watch_tele_weekly_m, color = region)) +
  geom_point(show.legend = FALSE)+
  labs(x = "Read Newspaper Weekly", y = "Watch Television Weekly")
rwm/rnm/nwm

```

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References